---
title: "Analyzing geniusr with R"
author: "Your Name"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: html_document
---

### Vignette: Analyzing Drake's "If You're Reading This It's Too Late" with `geniusr`

#### Setup

First, ensure you have `geniusr` installed and loaded. You might also need a Genius API token.

```{r}
# Install geniusr if you haven't already
if (!requireNamespace("geniusr", quietly = TRUE)) {
  install.packages("geniusr")
}

# Load geniusr
library(geniusr)
library(tidyverse)


genius_api_token <- Sys.getenv("genius_api_token")
# Set your Genius API token
genius_token(genius_api_token)
```

#### Fetching the Album Data

To analyze the album, we first need to fetch its tracks and lyrics. The `geniusr` package offers functions to search for artists, albums, and song lyrics.

```{r}
# Search for Drake
drake <- search_artists("Drake")

# Get album information
album_info <- get_albums(artist_id = drake$id[1], search = "If You're Reading This It's Too Late")

# Fetch tracks of the album
album_tracks <- get_album_tracks(album_info$id[1])
```

#### Analyzing Lyrics

With the track information, we can proceed to fetch and analyze the lyrics. For this example, let's explore the frequency of specific words or themes across the album.

```{r}
# Function to fetch lyrics and count words
count_words_in_song <- function(song_id, word) {
  lyrics <- get_song_lyrics(song_id)
  sum(str_count(lyrics, ignore_case(word)))
}

# Example: Count occurrences of the word 'love'
word_counts <- sapply(album_tracks$id, count_words_in_song, word = "love")

# Display the count
names(word_counts) <- album_tracks$title
print(word_counts)
```

#### Visualization

Let's visualize the frequency of the word "love" across the tracks.

```{r}
library(ggplot2)

# Create a dataframe for plotting
df <- data.frame(track = names(word_counts), count = word_counts)

# Plot
ggplot(df, aes(x = reorder(track, -count), y = count)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Frequency of 'love' in 'If You're Reading This It's Too Late'",
       x = "Track",
       y = "Count")
```

#### Conclusion

This vignette demonstrated how to utilize the `geniusr` package to analyze the lyrics of Drake's album "If You're Reading This It's Too Late". By fetching album and track information, and then exploring the lyrics, we can uncover insights about thematic elements across the album.

---

Please, replace `"<your_genius_api_token>"` with your actual Genius API token. This vignette provides a basic framework for lyrical analysis which can be expanded based on specific analytical goals or interests.