---
title: "Census Data Retrieval for San Luis Obispo County, CA"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This vignette demonstrates how to retrieve decennial Census data for San Luis Obispo County, California using the `tidycensus` package in R.

## Step 1: Get API Key

First, you need to obtain an API key from the U.S. Census Bureau. You can sign up for a free API key at [Census Bureau API Key Signup](https://api.census.gov/data/key_signup.html).

```{r}
# Replace with your actual API key
api_key <- "25a51aebc8cc7e3b63afadedb7572521c320d6d6"
```

## Step 2: Install and Load Necessary Packages

Next, install and load the necessary packages. If the packages are already installed, this step will just load them.

```{r}
# Install necessary packages if not already installed
if (!require(tidycensus)) install.packages('tidycensus')
if (!require(ggplot2)) install.packages('ggplot2')
if (!require(terra)) install.packages('terra')
if (!require(r2r)) install.packages('r2r')

# Load the packages
library(tidycensus)
library(ggplot2)
library(terra)
library(r2r)
```

## Step 3: Set Census API Key

Set the Census API key. This step is necessary to authenticate your requests to the Census API.

```{r}
# Set Census API key
census_api_key(api_key, install=TRUE, overwrite=TRUE)
readRenviron("~/.Renviron")
```

## Step 4: Define Parameters

Define the parameters for the data retrieval. In this case, we're interested in data for San Luis Obispo County, California from the 2020 decennial Census.

```{r}
# Define county and state for San Luis Obispo, CA
county <- "079"  # FIPS code for San Luis Obispo County
state <- "06"    # FIPS code for California
year <- 2020     # Year of the decennial census
```

## Step 5: Define Variables

Define the variables you're interested in. These correspond to specific racial and ethnic categories.

```{r}
race_vars <- c(
  Hispanic = "P2_002N",
  White = "P2_005N",
  Black = "P2_006N",
  Asian = "P2_008N",
  AI_AN = "P2_007N", ## American Indian and Alaska Native
  NH_PI = "P2_009N", ## Native Hawaiian and Other Pacific Islander
  some_Other = "P2_010N", ## Other
  Two_or_more = "P2_011N" ## Two or more
)
```

## Step 6: Retrieve Data

Finally, retrieve the data using the `get_decennial` function from the `tidycensus` package.

```{r}
slo_ca_df <- get_decennial(geography = "county", 
              variables = race_vars, 
              state = state, 
              county = county, 
              year = year)
```

## Conclusion

You now have a data frame `slo_ca_df` that contains the decennial Census data for San Luis Obispo County, California. You can use this data for further analysis or visualization.
