---
title: "Web Scraping Sports Data from Sports Reference"
output: html_document
---

```{r setup, include=FALSE}
library(rvest)
library(dplyr)
```

## Introduction

This vignette demonstrates how to scrape sports data from Sports Reference. We will be using the `rvest` package to read and extract information from the website, and `dplyr` for data manipulation.

## Scraping General Stat Info
Explain the process of scraping general stat info from the website.





## Define important variables
  
```{r}
year <- 2019
type_info <- "passing"
```


```{r}
url <- paste0("https://www.pro-football-reference.com/years/",year,"/",type_info,".htm")
```

## Read the HTML content of the webpage
```{r}
webpage <- read_html(url)
```

## Extract the table from the webpage
```{r}
table <- html_node(webpage, "table")
```

## Convert the table to a data frame
```{r}
table_df <- html_table(table)
```

## Print the data frame
```{r}
table_df
```

## Conclusion

With this function, you can easily scrape and analyze sports data from Sports Reference. Happy data analysis!
