---
title: "Analyzing Spotify's Stock with tidyquant"
author: "Your Name"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: html_document
---



```{r setup, include=FALSE}
# Install and load the spotifyr package
if (!require(tidyquant)) {
  install.packages('tidyquant')
}
library(tidyquant)
```

### Introduction

In this document, we explore the capabilities of the `tidyquant` package to access and analyze financial data. Specifically, we'll focus on Spotify's stock (ticker symbol: SPOT) to demonstrate how to retrieve stock prices and perform a simple time series plot.

### Getting Spotify's Stock Data

The `tidyquant` package simplifies the process of fetching financial data. To get Spotify's stock data, we can use the `tq_get()` function:

```{r get-spotify-data}
spotify_stock <- tq_get("SPOT", from = "2020-01-01", to = "2020-12-31")
head(spotify_stock)
```

This code retrieves the daily stock prices for Spotify for the year 2020.

### Visualizing Stock Trends

Let's create a time series plot to visualize Spotify's stock price movement over 2020:

```{r plot-spotify-stock}
spotify_stock %>%
  ggplot(aes(x = date, y = adjusted)) +
  geom_line(color = 'blue') +
  labs(title = "Spotify Stock Price in 2020",
       x = "Date",
       y = "Adjusted Close Price ($)") +
  theme_minimal()
```

### Conclusion

The `tidyquant` package offers an efficient way to access and analyze financial data in R. By leveraging its functionality, we can easily visualize trends in Spotify's stock prices, facilitating better financial analysis and decision-making.

