---
title: "Exploring Spotify Data with spotifyr"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is a short demonstration of how to use the `spotifyr` package in R to interact with the Spotify API.

```{r}
# Install and load the spotifyr package
if (!require(spotifyr)) {
  install.packages('spotifyr')
}
library(spotifyr)

# Set Spotify API credentials
Sys.setenv(SPOTIFY_CLIENT_ID = 'your_client_id')
Sys.setenv(SPOTIFY_CLIENT_SECRET = 'your_client_secret')

# Authenticate with Spotify API
access_token <- get_spotify_access_token()

# Search for a track
track <- get_track_audio_features('Never Gonna Give You Up')

# Print track features
print(track)
```

In this example, we authenticate with the Spotify API, search for a specific track ('Never Gonna Give You Up'), and print its audio features. You can replace the track name with any other track to get its features.
```

Please note that the `spotifyr` package requires you to have a Spotify Developer account and to set up an application to get the necessary client ID and client secret. Also, the `get_track_audio_features` function may not work if the track is not available in the Spotify catalog in your region. You can find more information about the `spotifyr` package and its functions in its [documentation](https://www.rdocumentation.org/packages/spotifyr/versions/2.2.3).